---
title: "P_cycling_bacteria"
author: "MD"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Load libraries
```{r}
library(vegan)
library(pairwiseAdonis)
library(RColorBrewer)
library(microbiomeMarker)
library(tidyverse)
library(phyloseq)
library(readxl)
```


# Read in data
```{r}
kegg <- read_excel('Data/EMU_database.GIBBs.KO.PICRUST2.xlsx', sheet='EMU_database.GIBBs.PICRUSt2', range='B9:BE17565')
kegg <- kegg %>%
  mutate_if(is.numeric, replace_na, replace=0)
```

# Select p decomp
```{r}
wanted <- kegg[kegg$phoA > 0 | kegg$phoD > 0 | kegg$PHO > 0 | kegg$appA > 0 | kegg$phoN > 0 ,  ]$Species
P.sub <- subset_taxa(P.rel, Species %in% wanted)

P.glom <- tax_glom(P.sub, taxrank="Species")
P.glom <- prune_taxa(taxa_sums(P.glom) > 0, P.glom)
pdecomp <- psmelt(P.glom) %>% 
  filter(., Abundance != 0)

```

# Run model (log transform needed)
```{r}
mod <- lm(log(Abundance) ~  Stage*Fertilization*Domestication, data=pdecomp)
par(mfrow=c(2,2))
plot(mod)
summary(mod)
```

# Subset data for each time point and fertilization
```{r}
ds <- filter(pdecomp, Stage == 'Seedling')
dev <- filter(pdecomp, Stage == 'Early_Vegetative')
dlv <-filter(pdecomp, Stage == 'Late_Vegetative')
def <- filter(pdecomp, Stage == 'Early_Flower')
dlf <- filter(pdecomp, Stage == 'Late_Flower')
ds_f <- filter(ds, Fertilization == 'Fertilized')
ds_u <- filter(ds, Fertilization == 'Unfertilized')
dev_f <- filter(dev, Fertilization == 'Fertilized')
dev_u <- filter(dev, Fertilization == 'Unfertilized')
dlv_f <- filter(dlv, Fertilization == 'Fertilized')
dlv_u <- filter(dlv, Fertilization == 'Unfertilized')
def_f <- filter(def, Fertilization == 'Fertilized')
def_u <- filter(def, Fertilization == 'Unfertilized')
dlf_f <- filter(dlf, Fertilization == 'Fertilized')
dlf_u <- filter(dlf, Fertilization == 'Unfertilized')
```

# KW tests
```{r}
kruskal.test(Abundance ~ Domestication, data = ds_f)  # ***
kruskal.test(Abundance ~ Domestication, data = ds_u)  # NS
kruskal.test(Abundance ~ Domestication, data = dev_f) # ***
kruskal.test(Abundance ~ Domestication, data = dev_u) # NS
kruskal.test(Abundance ~ Domestication, data = dlv_f) # NS
kruskal.test(Abundance ~ Domestication, data = dlv_u) # NS
kruskal.test(Abundance ~ Domestication, data = def_f) # NS
kruskal.test(Abundance ~ Domestication, data = def_u) # ***
kruskal.test(Abundance ~ Domestication, data = dlf_f) # NS
kruskal.test(Abundance ~ Domestication, data = dlf_u) # ***
```

# Get data ready for graphs 
```{r}
pdec.sum <- summarize(group_by(pdecomp, Stage, Fertilization, Domestication),
                      n=n(),
                      mean=mean(Abundance),
                      sd=sd(Abundance),
                      se=sd/sqrt(n))
level_order <- factor(pdecomp$Stage, levels = c('Seedling','Early_Vegetative','Late_Vegetative','Early_Flower','Late_Flower'))
level_order.sum <- factor(pdec.sum$Stage, levels = c('Seedling','Early_Vegetative','Late_Vegetative','Early_Flower','Late_Flower'))
```

# Graph p decomp

```{r}
p <- ggplot() +
  geom_line(aes(color=Domestication,group=Domestication,
                x=level_order, y=Abundance), 
            data = pdecomp,
            stat='summary',fun='mean', size=1)+
  geom_errorbar(data=pdec.sum, width=0.25,alpha=0.5,
                aes(ymin = mean - se, ymax = mean + se, 
                    group=Domestication,color=Domestication,
                    x=level_order.sum))+
  scale_color_manual(values = c('darkblue','darkorange2'))+
  facet_wrap(~Fertilization, ncol=2)+
  theme_classic()+
  labs(y="Phosphorus Decomposer Relative Abundance")+
  theme(legend.title = element_blank(),
        legend.position = 'right',
        axis.text = element_text(color='black'),
        axis.title.x = element_blank());p
```


# Select P solubilizer
```{r}
wanted <- kegg[kegg$pqqC > 0 ,  ]$Species
P.sub <- subset_taxa(P.rel, Species %in% wanted)

P.glom <- tax_glom(P.sub, taxrank="Species")
P.glom <- prune_taxa(taxa_sums(P.glom) > 0, P.glom)
psol <- psmelt(P.glom) %>% 
  filter(., Abundance != 0)
```

# Subset for each time point and fertilization
```{r}
ds <- filter(psol, Stage == 'Seedling')
dev <- filter(psol, Stage == 'Early_Vegetative')
dlv <-filter(psol, Stage == 'Late_Vegetative')
def <- filter(psol, Stage == 'Early_Flower')
dlf <- filter(psol, Stage == 'Late_Flower')
ds_f <- filter(ds, Fertilization == 'Fertilized')
ds_u <- filter(ds, Fertilization == 'Unfertilized')
dev_f <- filter(dev, Fertilization == 'Fertilized')
dev_u <- filter(dev, Fertilization == 'Unfertilized')
dlv_f <- filter(dlv, Fertilization == 'Fertilized')
dlv_u <- filter(dlv, Fertilization == 'Unfertilized')
def_f <- filter(def, Fertilization == 'Fertilized')
def_u <- filter(def, Fertilization == 'Unfertilized')
dlf_f <- filter(dlf, Fertilization == 'Fertilized')
dlf_u <- filter(dlf, Fertilization == 'Unfertilized')
```

# Kruskal wallis test for each time point and fertilization
```{r}
kruskal.test(Abundance ~ Domestication, data = ds_f)  #NS
kruskal.test(Abundance ~ Domestication, data = ds_u)  #NS
kruskal.test(Abundance ~ Domestication, data = dev_f) #***
kruskal.test(Abundance ~ Domestication, data = dev_u) #NS
kruskal.test(Abundance ~ Domestication, data = dlv_f) #NS
kruskal.test(Abundance ~ Domestication, data = dlv_u) #NS
kruskal.test(Abundance ~ Domestication, data = def_f) #NS
kruskal.test(Abundance ~ Domestication, data = def_u) #***
kruskal.test(Abundance ~ Domestication, data = dlf_f) #NS
kruskal.test(Abundance ~ Domestication, data = dlf_u) #***
```

# Create df for graph (summary for error bars and factor levels for x axis)
```{r}
psol.sum <- summarize(group_by(psol, Stage, Fertilization, Domestication),
                      n=n(),
                      mean=mean(Abundance),
                      sd=sd(Abundance),
                      se=sd/sqrt(n))
level_order <- factor(psol$Stage, levels = c('Seedling','Early_Vegetative','Late_Vegetative','Early_Flower','Late_Flower'))
level_order.sum <- factor(psol.sum$Stage, levels = c('Seedling','Early_Vegetative','Late_Vegetative','Early_Flower','Late_Flower'))

```

# Graph p solubilizers
```{r}
p <- ggplot() +
  geom_line(aes(color=Domestication,group=Domestication,
                x=level_order, y=Abundance), 
            data = psol,
            stat='summary',fun='mean', size=1)+
  geom_errorbar(data=psol.sum, width=0.25,alpha=0.5,
                aes(ymin = mean - se, ymax = mean + se, 
                                  group=Domestication,color=Domestication,
                                  x=level_order.sum))+
  scale_color_manual(values = c('darkblue','darkorange2'))+
  facet_wrap(~Fertilization, ncol=2)+
  theme_classic()+
  labs(y="Phosphorus Solubilizer Relative Abundance")+
  theme(legend.title = element_blank(),
        legend.position = 'right',
        axis.text = element_text(color='black'),
        axis.title.x = element_blank());p
```

