---
title: "qPCR"
author: "MD"
date: '`r Sys.Date()`'
output: 
 html_document:
   toc: true
   toc_float: true
   toc_level: 3
   editor_options:
     chunk_output_type: console
---



```{r, include = F}
library(tidyverse)
library(car)
library(emmeans)
library(agricolae)
```

# Read in data 

I copied the sheet in the excel workbook with the qPCR data and saved that as a csv. Here, I read it in and corerce the factors to be in an established order (as opposed to alphabetical). 

```{r read in data, results='hide'}
d <- read.csv("Data/qPCR.csv") %>% 
  filter(!is.na(Concentration..16S.copies.g.soil.DW.))

d <- d %>%
mutate(Fertilization = as_factor(Fertilization)) %>%
mutate(Fertilization = fct_relevel(Fertilization, "Unfertilized", "Fertilized"))
table(d$Fertilization)

d <- d %>%
mutate(Harvest = as_factor(Harvest)) %>%
mutate(Harvest = fct_relevel(Harvest, "Seedling", "Early_Vegetative", "Late_Vegetative", "Early_Flower", "Late_Flower"))
table(d$Harvest)

options(contrasts = c("contr.sum", "contr.poly"))
par(mfrow=c(2,2))
```

# Model fitting

## Untransformed data

First, I ran the linear model using untransformed data. It looks like we can proceed with untransformed data.
```{r, no transform model}
mod <- lm((new.16s) ~  Fertilization*Domestication*Harvest, 
           data = d)
plot(mod) 
```


## ANOVA

```{r, anova}
summary(aov(mod))
  
```

We can see from the ANOVA table that we have a significant main effect of domestication. We do not have a main effect of domestication.

We also see  interactive effects occurring. 

### Main effect examinination

We can first examine the main effects:

```{r, main effect fertilization}
t<-HSD.test(mod, "Domestication");t
```
Wild tomato soils have greater microbial biomass than modern.


### Interactive effects

To examine interactive effects, I like to first look at an estimated marginal means plot. This plot gives us an idea of what patterns are occurring. 

```{r, emmip, warning=F}
p <- emmip(mod, Domestication ~ Harvest | Fertilization, CIs = TRUE);p
```


From this graph, it seems like tentatively, fertilized wild > modern at the early flower stage. It also looks like there are two stages in unfertilized where wild has greater microbial biomass. 


This below code tells me where to put the stars to indicate what pairs are significant. The output is long, so I will only list the code here. The results from this line of code indicate that at Late_Flower and Early_Flower in the P- condition, Wild > Modern. In P+, early flower shows differences. There were no other significant pairings. 
```{r, pairwise contrasts, results='hide'}
emmeans(mod, pairwise ~ Domestication|Harvest|Fertilization)
```

Now, we create our dataframe with the back-transformed data and the letters denoting differences
```{r, df generation}
df <- as.data.frame(emmeans(mod, ~ Fertilization*Domestication*Harvest))

res <- emmeans(mod, list(pairwise ~ Fertilization*Domestication*Harvest), adjust = "fdr")
groups <- data.frame(multcomp::cld(res[[1]], adjust='fdr', Letters=c("abcdefghijklmnopqrstuvwxy"), decreasing=T)) %>% 
  select(., Fertilization, Domestication, Harvest, .group)
df <- inner_join(df,groups, by = c('Fertilization', 'Domestication', 'Harvest'))

names(df)[names(df) == '.group'] <- 'letters'
df$letters <- gsub(' ','',df$letters)
```

# Visualization

Now, we can visualize our results. Here is a line graph showing the P+ and P- conditions. 

```{r, plot, warning=F}
dff <- filter(df, Fertilization == "Fertilized")
dfu <- filter(df, Fertilization == "Unfertilized")
p_f <- ggplot(aes(x=Harvest, y=emmean), data = dff) +
  geom_line(aes(color=Domestication, group=Domestication),
            stat='summary',fun='mean',linewidth=1) +
  geom_errorbar(width=0.25,alpha=0.5,
                aes(ymin = lower.CL, ymax = upper.CL, 
                    group=Domestication,color=Domestication))+
  scale_color_manual(values = c('darkblue','darkorange2'))+
  geom_text(aes(label = letters),position=position_dodge(0.9), vjust = -3, hjust = -0.2) +
  facet_wrap(~Fertilization, ncol=2)+
  theme_classic()+
   ylab(expression(atop("Bacterial Concentration", paste("(16S copies /g soil DW)")))) +
  theme(legend.title = element_blank(),
        legend.position = 'right',
        strip.text = element_blank(),
        axis.text = element_text(color='black'),
        axis.title.x = element_blank());p_f

#ggsave(plot = p_f, "../../Manuscript/figs/R_output/qpcr.png", height = 4, width = 8)
```

